---
title: "Introduction to the electionsBR package"
author: "Fernando Meireles, Denisson Silva, Beatriz Costa"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to electionsBR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo=FALSE}
library(electionsBR)
```


## Overview

`electionsBR` is an `R` package used to pull and clean Brazilian electoral data, directly from the Superior Electoral Court (TSE) [website](http://www.tse.jus.br/eleicoes/estatisticas/repositorio-de-dados-eleitorais). Among others, `electionsBR` retrieves data on electoral results, candidates' personal and professional backgrounds, parties' electoral performances, electoral coalitions, available seats under dispute, and voters' profile.

## How to install

Since the version `0.1.1`, `electionsBR` is on CRAN and can be installed with:

```{r, eval=FALSE}
install.packages("electionsBR")
```

You can install these pre-release versions from [GitHub](https://github.com/) with:

```{r, eval=FALSE}
if (!require("devtools")) install.packages("devtools")
devtools::install_github("silvadenisson/electionsBR")
```

## How to use

No previous experience with `R` is required to use the `electionsBR` package. In fact, it takes only two lines of code to download, clean and export (to [Stata](https://www.stata.com/) and [SPSS](https://www.ibm.com/analytics/us/en/technology/spss/) formats) Brazilian electoral data. For example, here is how we get a full, cleaned dataset with electoral results for the 2010 Federal election:

```{r, eval=FALSE}
install.packages("electionsBR")
electionsBR::vote_mun_zone_fed(2010, export = TRUE)
```

By setting the `export` argument to `TRUE` in the `vote_mun_zone_fed` function, the package will download and clean the relevant data directly from the TSE website and save it in the `R` working directory (the function automatically tells the user where this directory is located) in two different files:

1. `electoral_data.dta`, to be used with [Stata](https://www.stata.com/);
2. `electoral_data.sav`, to be used with [SPSS](https://www.ibm.com/analytics/us/en/technology/spss/).

## Data on elections results

`electionsBR`'s chief functions are the ones used to get electoral results, which includes candidates' number of votes and electoral status (i.e., elected and non-elected), among many others. To get these data, use `vote_mun_zone_` functions: 

```{r, eval=FALSE}
# Load the package
library(electionsBR)

# Use vote_mun_zone_fed function to download 2014 election results
df <- vote_mun_zone_fed(2014)
```

```{r, eval=FALSE}
# Get results for the 2000 local elections
df <- vote_mun_zone_local(2000)
```

## Federal elections results

There are also functions used to quickly access presidential and legislative elections results per state and per municipality -- with vote proportions (i.e., vote share) reported. To get presidential runoff (second round) results in runoff, use `president_` functions:

```{r, eval=FALSE}
# Get presidential electoral results by municipality in tidy format
pres_mun <- president_mun_vote(2010, prop = TRUE)

# Get presidential electoral results by state in tidy format
pres_st <- president_state_vote(2010, prop = TRUE)
```

To get legislative election results, use instead:

```
legislative_mun_vote(year)
legislative_state_vote(year)
```

## Data on running candidates

Data on candidates' party affiliation, ballot number, backgrounds and personal information can be retrieved with `candidate_` functions:

```{r, eval=FALSE}
# Get data on candidates running in local elections
df <- candidate_local(2016)

# Get data on candidates running in federal elections
df <- candidate_fed(2016)
```

## Data on parties and electoral coalitions

Parties' electoral performances in both local and federal elections, for all electoral positions, are extracted with:

```{r, eval=FALSE}
# Get data on parties' electoral performances in the federal election of 2006
federal <- party_mun_zone_fed(2006)

# Get data on parties' electoral performances in the local election of 2004
municipal <- party_mun_zone_local(2004)
```

To get the composition of electoral coalitions (common in proportional elections), use instead:

```
legend_fed(year)
legend_local(year)
```

## Data on voters

To get data on voters' profile in both local and federal elections, use the `voter_profile` function:

```{r, eval=FALSE}
# Get data on voters' profile
voters02 <- voter_profile(2002)
voters04 <- voter_profile(2004)
```

Individual records of partisan affiliation are extracted with the `voter_affiliation` function:

```{r, eval=FALSE}
# PT (Brazilian Workers' Party) in Distrito Federal (DF)
df <- voter_affiliation("PT", "DF")

# PSDB (Brazilian Social Democracy Party) in Sao Paulo (SP)
df <- voter_affiliation("PSDB", "SP")

# PT and PC do B (Brazilian Communist Party) in DF, MG and AL
df <- voter_affiliation(c("PT", "PC do B"), c("DF", "MG", "AL"))
```

To consult state abbreviations, use:

```{r}
uf_br()
```

A `character` vector with a list os party abbreviations, valid for 2016 elections, can be obtained with:

```
parties_br(year)
```

## Other functions

`electionsBR` also includes functions to download data on valid and residual votes and on the number of seats under dispute in each election. The relevant functions are, respectively:

```
details_mun_zone_fed(year)
details_mun_zone_local(year)
```

```
seats_fed(year)
seats_local(year)
```

## Optional arguments

### Exporting Brazilian electoral data

Most `electionsBR`'s functions accept an `export` argument (`logical`, must be `TRUE` or `FALSE`; defaults to the latter) controlling whether the functions should export the retrieved data to [Stata](https://www.stata.com/) and [SPSS](https://www.ibm.com/analytics/us/en/technology/spss/) files or not.

```{r, eval=FALSE}
df <- party_mun_zone_fed(2010, export = TRUE)
```

### Removing UTF-8 special characters from texts

By default, `electionsBR`'s functions maintain data's original encoding (`Latin-1`) in special characters. To convert strings to `ASCII`, set the `ascii` argument to `TRUE`.

```{r, eval=FALSE}
df <- candidate_fed(2010, ascii = TRUE)
```

In Mac computers, this option may cause errors (or may retrieve incomplete data for the 2016 elections). To avoid them, you may use the `encoding` optional argument as follows:

```{r, eval=FALSE}
df <- candidate_fed(2010, ascii = TRUE, encoding = "Latin-1")
```

`encoding` may also be `UTF-8` or other valid encodings.

### Filtering results by state

Sometimes, getting state electoral data, and not for the whole country, is what one needs. To achieve this, use the `uf` optional argument (available in most functions):

```{r, eval=FALSE}
# Electoral results for the 2010 federal elections in Sao Paulo (SP)
df <- vote_mun_zone_fed(2010, uf = "SP")

# Electoral results for the 2010 federal elections in Minas Gerais (MS)
df <- vote_mun_zone_fed(2010, uf = "mg")

# Electoral results for the 2010 federal elections in RS, SC, and PR
df <- vote_mun_zone_fed(2010, uf = c("RS", "SC", "PR"))
```

Notice that the input must be a `character` vector -- with case insensitive state abbreviations (MG, Mg, mG, and mg are all equally valid inputs).

## A complete list of functions

* `candidate_fed()`
* `candidate_local()`
* `details_mun_zone_fed()`
* `details_mun_zone_local()`
* `legend_fed()`
* `legend_local()`
* `legislative_mun_vote()`
* `legislative_state_vote()`
* `parties_br()`
* `party_mun_zone_fed()`
* `party_mun_zone_local()`
* `president_mun_vote()`
* `president_state_vote()`
* `seats_fed()`
* `seats_local()`
* `uf_br()`
* `vote_mun_zone_fed()`
* `vote_mun_zone_local()`
* `voter_affiliation()`
* `voter_profile()`


## Elections in Brazil

All the data retrieved by `electionsBR` is made available online by the [Brazilian Superior Electoral Court (TSE)](http://english.tse.jus.br/the-electoral-justice/the-electoral-justice-1/the-electoral-justice). As state in the [English version of the TSE website](http://english.tse.jus.br/the-brazilian-electoral-system/elections/elections), elections in Brazil are organized in the following way:

> Are carried out together the federal elections and the state elections, for president and vice president, senator and alternates, governor and vice governor, federal deputies and state deputies. Two years later, will be municipal elections, of mayor and vice mayor and city councils. Note that the distinguishing criterion is by federal level and not by powers -- the representatives from the executive and legislative are elected at the same time. 

> In case of elections for president and vice president, governor and vice governor, senator and alternates and mayor and vice mayor, the system used is the majoritary, in which the most voted candidate will be elected. If none of candidates for president, governor or mayor from a municipality over 200,000 habitants reach the absolute majority of the valid votes, will be convened the second round with the two most voted.

> Now for the city council elections and state deputy and federal deputy, the criterion is proportional, which considers not only the candidate votes, but also his party’s votes. Hence, not always the best voted candidate will be elected. The filling of vacancies will depend on the performance of the whole group of candidates of the party or alliance.


### Official documentation

`electionsBR` internal documentation is entirely based on the official documentation provided by the TSE in the [Repositorio de Dados Eleitorais](http://www.tse.jus.br/eleicoes/estatisticas/repositorio-de-dados-eleitorais).^[To access the documentation for each type of electoral data, it is necessary to download and open the correspondent compressed files (`.zip`).]

## Disclaimer

The `electionsBR` package does not modify nor filter the data provided by the TSE -- despite its many problems.^[The only exceptions are the `president_` and `legislative_` functions, which aggregates electoral data in a different unit than in the original files.] Additionally, users must be aware that the TSE updates its databases frequently, so it is important to register when a version of the electoral data is used. In any case, we are not responsible for problems with the data users may find.

## How to cite

To cite `electionsBR` in publications, please use:

``` {r}
citation("electionsBR")
```

