---
title: "Introduction to the electionsBR package"
author: "Vignette Author"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to electionsBR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Overview

`electionsBR` is an `R` package used to automatically retrieve and clean electoral data from Brazil, directly from the Superior Electoral Court (TSE) [website](http://www.tse.jus.br/eleicoes/estatisticas/repositorio-de-dados-eleitorais). Among others, `electionsBR` provides functions to obtain data on electoral results, candidates' personal and professional backgrounds, parties' electoral performances, electoral coalitions, available seats under dispute, and voters' profile.

## How to install

Since the version `0.1.1`, `electionsBR` is on CRAN and can be installed with:

```{r, eval=FALSE}
install.packages("electionsBR")
```

Development versions of the package are also available on [GitHub](https://github.com/). You can install these pre-release versions with:

```{r, eval=FALSE}
if (!require("devtools")) install.packages("devtools")
devtools::install_github("silvadenisson/electionsBR")
```

## Usage

No previous experience with `R` is required to use the `electionsBR` package. In fact, one needs only two lines of code to download, clean and export (to Stata and SPSS formats) Brazilian electoral data. For instance, here is how we get a full, cleaned dataset with electoral results for the 2010 Federal election:

```{r, eval=FALSE}
install.packages("electionsBR")
electionsBR::vote_mun_zone_fed(2010, export = TRUE)
```

By setting the `export` argument to `TRUE` in the `vote_mun_zone_fed` function, the package will download and clean the relevant data directly from the TSE website and save it in the `R` working directory (the function automatically tells the user where this directory is located) in two different files:

1. `electoral_data.dta`, to be used with [Stata](http://www.stata.com/);
2. `electoral_data.sav`, to be used with [SPSS](https://www.ibm.com/analytics/us/en/technology/spss/).

In the remaining of this vignette, we document all the functionality included in the package -- the table below sums up this information.

## Data on electoral results

```{r, eval=FALSE}
# Load the package
library(electionsBR)

# Use vote_mun_zone_fed function to download 2014 electoral results
df <- vote_mun_zone_fed(2014)
```

```{r, eval=FALSE}
# Get results for the 2000 local election
df <- vote_mun_zone_local(2000)
```

## Data on running candidates





## Data on parties and electoral coalitions

## Data on voters

## A complete list of functions

## Other options

### Exporting Brazilian electoral data

### Removing UTF-8 special characters from texts

### Filtering results by state

## An easier way to get Federal electoral results

## Brazilian Electoral Supreme Court (TSE)

### Official documentation

## Citation
